### 2011-10-21 doc/manuscript.tex
Just an update.

% Four potential reviewers whose e-mail addresses you can provide
% Claus O. Wilke
% Debra Bessen
% Bruce R. Levin
% Michael Wessels

@@

Molecular evolution of bacterial genomes is multifaceted. Studies of bacterial
evolution often focused on one side of many evolutionary processes.  While
understanding one aspect of bacterial evolutionary processes we could be
ignorant of the other features.  Horizontal gene transfer process was often
considered in macro-evolutionary scale using a great deal of bacterial species.
As closely related species or groups of bacterial individuals are becoming a
focus of evolutionary study, recombining gene transfer process would surface as
another facet of evolutionary forces.  Because the two gene transfer processes
would act on closely related bacterial genomes simultaneously, ignoring one of the two processes
in studying evolution of bacteria would lead to a narrowed point of view to the
underlying evolutionary forces. Our study of SPY and SDE genomes undertook the
work of discriminating recombining and horizontal gene transfer
processes because the two species were very closely related.  We showed that the
two gene transfer processes acted on different biological functions. 
Recombining
gene transfer process in the analysis of SPY and SDE genomes was related with
genes encoding membrane-bound molecules and aminoacyl tRNA synthetases. On the
contrary, horizontal gene transfer process was found to be associated with
transposases, which would allow horizontal gene transfer of their neighboring
genes.  Because SPY was a human-specific pathogen, and SDE was an opportunistic
human pathogen, we were interested in association of pathogenicity with the two
gene transfer processes. We showed that virulence genes were more associated
with 
horizontal gene transfer process than with recombining gene transfer process.
With the current data of only SPY and SDE genomes we could not conclude that the
association of virulence genes with horizontal gene transfer not recombining
gene transfer was applicable to other species.  However, without the
discrimination of the two processes it would have been difficult to find the
disparate association of biological functions or virulence genes with the
different gene transfer processes.

Bacterial gene transfer processes as an evolutionary force would affect more
frequently in some biological functions than in others.  In our attempt of
finding which biological functions were associated with recombining or
horizontal gene transfers, we devised a measure of recombination intensity for
genes by using frequency of recombinant edges that affect those genes, and used
frequency of horizontal gene transfer events.  This approach allowed us to be
able to associate gene transfer with biological functions including
membrane-bound molecules, aminoacyl tRNA synthetases, and ribosomes.  Sugar
phosphotransferases could be important in signal transduction to recognize
environmental changes.  Transporters, and fatty acid and lipid biosyntheses
would be involved in membrane-bound molecules.  Because bacteria as a
unicellular organism should be able to respond to changes to environment, some
evolutionary pressures must have been acting on genes encoding membrane-bound
molecules. Therefoure, recombining gene transfer process appeared to be more
associated with biological functions that had been critical of sensing queues of
bacterial environmental changes.  Because several aminoacyl tRNA synthetases
belonging to gene ontology term of translation (GO:0006412) had been a focus of
gene transfer in bacterial species \citep{Woese2000}, we could confirm that the
recombination intensity and requency of recombining gene transfer events would
be useful in finding genes associated with recombining gene transfer.
The biological role of transposases was concordant with horizontal gene transfer
process because the process involved in transposases would allow bacteria to
acquire foreign genes without the requirement of homologous regions in the
bacteria.  These differences in the profile of functional categories association
of the two gene transfer processes were because of the approach of dicriminating
the two gene transfer processes.

Functional categories concerning ribosomes were unexpectedly found to be
associated with recombining gene transfer.  Because genes related with ribosomes
were considered to be housekeeping genes, gene transfer in which would be rare,
it was unexpected that ribosome related genes were found to be associated with
recombining gene transfer.  Because it was a bizarre result, we investigated the
association by considering different recombinant edge types.  Ribosome related
functional categories were found to be top ranked if we took
tree-topology-preserving recombinant edges into account (Tables
S\ref{tab:functional-notopology} and S\ref{tab:functional-all}).  whereas
ribosome related functional categories were not top ranked in Table
\ref{tab:functional}.  Therefore, high recombination intensity in ribosome
related genes were likely to be attributed to topology-preserving recombinant
edges that were found to be difficult to infer reliably. So, we would have to be
cautious of associating ribosome related functions with recombining gene
transfer.

Another contribution of our study to bacterial gene transfer research is to
elucidate the net directionality of gene flow from SPY to SDE in the core
genomes. Although the issue of net directionality has been tackled, our study
was the first attempt of resolving the issue in genome-wide scale. Yet, it still
remains to be seen whether this net directionality of gene flow from SPY to SDE
happened in the pan genome. A model-based method with more refined models would
be necessary for the reserach direction because the parsimony-based method we
employed simply could not show any net directionality between SPY and SDE in the
pan genome. Although we are unsure whether much larger number of genomes could
be useful in finding net directionality, we consider that more SDE genomes would
be beneficial to future studies.

Were there unbalanced recombining gene transfer between the two species, what
were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
the species?  Kalia et al.\ \citet{Kalia2001} discussed restriction-modification system (RMS)
as a possible mechanism of the unbalanced gene flow.  Restriction-modification
system is a biological self-defense mechanism of bacteria, which is diverse in
prokaryote world.  Biological roles of RMS might be to maintain and control
species identity \citep{Jeltsch2003}.  Recently, Budroni et al.\ \citet{Budroni2011a} studied
that \textit{Neisseria meningitidis} phylogenetic clades were associated with
RMS, claiming that RMS modulated homologous recombination in the bacteria of
\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contained
restriction-modification type I system as well as type II system. The two
genomes of SDE contained only type II system. More tight self-defense system of
SPY may have made the species less capable of aquiring foreign genetic material
from SDE, and accordingly relatively unbalanced gene flow between SPY and SDE
may have been associated with different restriction-modification systems in the
two species \citep{Kalia2001}.

We reexamined the seven genes that Kalia et al.\ \citet{Kalia2001} had studied for net
directionality of gene transfer between SPY and SDE.  The internal fragments of
the six housekeeping genes (\textit{gki}, \textit{gtr}, \textit{murI},
\textit{mutS}, \textit{recP}, \textit{xpt}) were located in the core genome that
we investigated in this study.  Except for \textit{mutS} where Kalia et al.\ \citet{Kalia2001}
did not find recombination, our results were consistent with that of
Kalia et al.\ \citet{Kalia2001}.  Pinho et al.\ \citet{Pinho2010} reported recombination in the gene
\textit{parC}, in which we also observed moderate recombination.

A full model-based approach would be desirable in analyzing data for testing
relevant hypotheses more systematically. Model complexity might not allow easy
access to the implementation of a model. Our study involved a model-based
approach and a parsimony-based one for understanding evolution in the core
genomes of several bacteria individuals. Because of inability of the model-based
method for studying the pan genomes we had to resort to the parsimony-based
method. However, the parsimony-based method was also applied to the core
genomes. We wished to compare the two methods in their results of the core
genome analyses.  Because of two independent analyses on the same data set, this
would allow us to be more confident of the results if the two analyses performed
similarly Generally, phylogenetic methods of detecting gene transfers often take
advantage of discrepancy between a reference species tree and gene trees. The
two approaches that we employed in this study were no exception.

It remains to be seen whether the inability of detecting net directionality
between SPY and SDE in the pan genomes was because of lack of power of the
parsimony-based approach that we employed. Considering gene duplication, loss,
and horizontal gene transfer in a model-based approach would be desirable in the
coming ages of abundant bacterial genomes. The model might consider genome
alignments where each taxon does not necessarily have a unique gene; some taxa
might have no gene due to gene loss, and other taxa multiple genes due to gene
gain, duplication, or horizontal gene transfer. With these types of models will
we be able to address more interesting evolutionary questions in bacteria. This
need of more refined models would be also true even for the study of net
directionality between species in the core genome.  In other words, one could
envision a model where a population two-taxa (i.e., SPY and SD groups) tree is
superimposed on the species tree shown in Figure S\ref{fig:clonalorigin}A. A
splitting event of the population tree could happen somewhere between the root
of the species tree and the internal node of SPY. Recombinant edges that connect
branches within SPY or SD clades would be more frequent than those that connect
branches between SPY and SDE clades.  With the reference to the recombination
within a clade one parameter can dictate the degree of gene from from SPY to
SDE, and another for the reverse direction.  If the two parameters could be
incorporated in the model, then they might be used to compare the degrees of
gene flow between the two species in a more statistically sound way.  We hope
that we will see thie refinement of the model to better illustrate bacterial
evolution in forthcoming bacterial genomic era.

@@

\section{Supplementary Information}

\subsection{Summarization of posterior samples for local gene trees and recombination intensity}

@@

\subsection{Simulation 1 - three population parameters}

@@

\subsection{Simulation 2 - numbers of recombinant edges from prior}

@@

\subsection{Simulation 3 - number of recombinant edges from posterior}

>>>>

% Four potential reviewers whose e-mail addresses you can provide
% Debra Bessen
% Claus O. Wilke
% Didelot Xavier
% Bruce R. Levin
% Michael Wessels

@@

Molecular evolution of bacterial genomes is multifaceted. Researches of
bacterial evolution often focused on one side of many evolutionary processes.
While understanding one aspect of bacterial evolutionary processes we could be
ignorant of the other features.  Horizontal gene transfer process was often
considered in macro-evolutionary scale using a great deal of bacterial species.
As closely related species or groups of individual bacteria are becoming a focus
of evolutionary study, recombining gene transfer process would surface as
another facet of evolutionary forces.  Because the two gene transfer processes
would act on closely related bacterial genomes simultaneously, ignoring one of
the two processes in studying evolution of bacteria would lead to a narrowed
point of view to the underlying evolutionary forces. Our study of SPY and SDE
genomes undertook the work of discriminating recombining and horizontal gene
transfer processes because the two species were very closely related.  In this
study, we showed that the two gene transfer processes acted on different
biological functions.  Recombining gene transfer process in the analysis of SPY
and SDE genomes was related with genes encoding membrane-bound molecules and
aminoacyl tRNA synthetases. On the contrary, horizontal gene transfer process
was found to be associated with transposases, which would allow horizontal gene
transfer of their neighboring genes.  Because SPY was a human-specific pathogen,
and SDE was an opportunistic human pathogen, we were interested in association
of pathogenicity with the two gene transfer processes. We also showed that
virulence genes were more associated with horizontal gene transfer process than
with recombining gene transfer process.  With the current data of only SPY and
SDE genomes we could not conclude whether the association of virulence genes
with horizontal gene transfer not recombining gene transfer was applicable
generally to cases of other species.  However, without the discrimination of the
two processes it would have been difficult to find the disparate association of
biological functions or virulence genes with the different gene transfer
processes.

We showed differences in the profile of functional category association of the
two gene transfer processes thanks to the approach of dicriminating the two gene
transfer processes.  Sugar phosphotransferases could be important in signal
transduction to recognize environmental changes.  Transporters, and fatty acid
and lipid biosyntheses would be involved in membrane-bound molecules.  Because
bacteria as a unicellular organism should be able to promptly respond to changes
to environment, some evolutionary pressures must have been acting on genes
encoding membrane-bound molecules.  We also found that the recombination
intensity and requency of recombining gene transfer events would be useful in
associating several aminoacyl tRNA synthetases genes belonging to gene ontology
term of translation (GO:0006412) with recombining gene transfer. It had been a
focus of gene transfer in bacterial species \citep{Woese2000}.  The biological
role of transposases was concordant with horizontal gene transfer process
because the process involved in transposases would allow bacteria to acquire
foreign genes without the requirement of homologous regions in the bacteria.

Another contribution of our study to bacterial gene transfer research is to
elucidate the net directionality of gene flow from SPY to SDE in the core
genomes. Although the issue of net directionality has been tackled, our study
was the first attempt of resolving the issue in genome-wide scale. Yet, it still
remains to be seen whether this net directionality of gene flow from SPY to SDE
happened in the pan genome. A model-based method with more refined models would
be necessary because the parsimony-based method we employed simply could not
show any net directionality between SPY and SDE in the pan genome. Although we
are unsure whether much larger number of genomes could be useful in finding net
directionality, more SDE genomes might be beneficial to future studies.

Were there unbalanced recombining gene transfer between the two species, what
were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
the species?  Kalia et al.\ \citet{Kalia2001} discussed restriction-modification
system (RMS) as a possible mechanism of the unbalanced gene flow.
Restriction-modification system is a biological self-defense mechanism of
bacteria, which is diverse in prokaryote world.  As biological roles of RMS
might be to maintain and control species identity \citep{Jeltsch2003}, Budroni
et al.\ \citet{Budroni2011a} recently studied that \textit{Neisseria
meningitidis} phylogenetic clades were associated with RMS, claiming that RMS
modulated homologous recombination in the bacteria of \textit{N.\ meningitidis}.
Most of \textit{S.\ pyogenes} genomes contained restriction-modification type I
system as well as type II system. The two genomes of SDE contained only type II
system. More tight self-defense system of SPY may have made the species less
capable of aquiring foreign genetic material from SDE \citep{Kalia2001}.

We reexamined the seven genes that Kalia et al.\ \citet{Kalia2001} had studied
for net directionality of gene transfer between SPY and SDE.  The internal
fragments of the six housekeeping genes (\textit{gki}, \textit{gtr},
\textit{murI}, \textit{mutS}, \textit{recP}, \textit{xpt}) were located in the
core genome that we investigated in this study.  Except for \textit{mutS} where
Kalia et al.\ \citet{Kalia2001} did not find recombination, our results were
consistent with those of Kalia et al.\ \citet{Kalia2001}.  Pinho et al.\
\citet{Pinho2010} reported recombination in the gene \textit{parC}, in which we
also observed moderate recombination.

It remains to be seen whether the inability of detecting net directionality
between SPY and SDE in the pan genomes was because of lack of power of the
parsimony-based approach that we employed. Considering gene duplication, loss,
and horizontal gene transfer in a model-based approach would be desirable in the
coming ages of abundant bacterial genomes. The model might consider genome
alignments where each taxon does not necessarily have a unique gene; some taxa
might have no gene due to gene loss, and other taxa multiple genes due to gene
gain, duplication, or horizontal gene transfer. With these types of models will
we be able to address more interesting evolutionary questions in bacteria. This
need of more refined models would be also true even for the study of net
directionality between species in the core genome.  In other words, one could
envision a model where a population two-taxa (i.e., SPY and SD groups) tree is
superimposed on the species tree shown in Figure S\ref{fig:clonalorigin}A. A
splitting event of the population tree could happen somewhere between the root
of the species tree and the internal node of SPY. Recombinant edges that connect
branches within SPY or SD clades would be more frequent than those that connect
branches between SPY and SDE clades.  With the reference to the recombination
within a clade one parameter can dictate the degree of gene from from SPY to
SDE, and another for the reverse direction.  If the two parameters could be
incorporated in the model, then they might be used to compare the degrees of
gene flow between the two species in a more statistically sound way.  We hope
that we will see this refinement of the model to better illustrate bacterial
evolution in forthcoming bacterial genomic era.

@@

\clearpage{}
\section*{Supplementary Information}

\subsection*{Summarization of posterior samples for local gene trees and recombination intensity}

@@

\subsection*{Simulation 1 - three population parameters}

@@

\subsection*{Simulation 2 - numbers of recombinant edges from prior}

@@

\subsection*{Simulation 3 - number of recombinant edges from posterior}
